Inference for multivariate peaks over threshold models

### 2021-11-11 evaluation.tex
updates

\subsection{Posterior predictive loss}
The posterior predictive loss criterion is introduced in \cite{gelfand1998}.  When we
  assume a squared error loss function, then for the $i$th observation, the posterior predictive
  loss criterion is computed as
  \begin{equation}
    \label{eq:ppl}
      S_{\text{PPL}}^k\left(P,x_i\right) = \text{Var}_P\left(X_i\right) + \frac{k}{k+1}\left(\text{E}_P\left[X_{i}\right] - x_i\right)^2
  \end{equation}
  where $X_i$ is a random variable from the posterior predictive distribution for $x_i$.  The
  scalar $k$ is a weighting factor that scales the importance of goodness of fit
  relative to precision.  In fact a small $\text{Var}(X_{i})$ indicates high
  precision, and a small $(\text{E}[X_{i}] - x_{i})^2$  indicates a good fit.  In our analysis, we take the limit as $k\to\infty$, weighting both
  equally. Overall, the smaller the score, the better.  Note that the PPLC score is defined for a univariate distribution.  As we are dealing
  with a multivariate distribution, we will be taking the posterior predictive loss
  summed over all dimensions.  That is,
  \begin{equation}
    \label{eq:ppl2}
      S_{\text{PPL}}\left(P,\bm{x}_i\right) = \frac{1}{d}\sum_{\ell = 1}^d\left[\text{Var}_P\left(X_{i\ell}\right) + \left(\text{E}_P\left[X_{i\ell}\right] - x_{i\ell}\right)^2\right] .
  \end{equation}

\subsection{Energy score}
We focus on \emph{energy scores} defined for a general probability distribution $P$, with finite expectation, as 
    \begin{equation}
    \label{eq:es}
    \text{S}_{\text{ES}}\left(P, \bm{x}_i\right) =  \text{E}_p g\left(\bm{X}_i, \bm{x}_i\right) -
                                    \frac{1}{2}\text{E}_p g\left(\bm{X}_i,\bm{X}_i^{\prime}\right),
    \end{equation}
where $g$ is a given kernel. The score defined in Equation \eqref{eq:es} can be evaluated using samples from $P$, with the help of the law of large numbers.
Moreover, Theorem~4 in \cite{gneiting2007}, states that if $g(\cdot,\cdot)$ is a negative definite kernel, then $\text{S}(P,\bm{x})$ is a \emph{proper} scoring rule.  Recall that a real valued function $g$ is a negative definite kernel if it is symmetric in its arguments, and $\sum_{i=1}^n\sum_{j=1}^na_ia_jg(x_i,x_j)\leq 0$ for all positive integers $n$, and any collection $a_1,\ldots,a_n\in{\mathbb R}$ such that  $\sum_{i = 1}^na_i = 0$.

In an Euclidean space, the conditions for a negative definite kernel are satisfied by the Euclidean distance. However, the Euclidean distance between two points will likely under-estimate the actual distance required to connect those points fully within ${\mathbb S}_{\infty}^{d-1}$.  Let
\begin{equation*}
    {\mathbb C}_{\ell}^{d-1} = \lbrace \bm{x} : \bm{x} \in {\mathbb S}_{\infty}^{d-1}, x_{\ell} = 1\rbrace
\end{equation*}
comprise the $\ell$th \emph{face}.  For points on the same face, Euclidean distance corresponds to the length of the shortest possible path in ${\mathbb S}_{\infty}^{d-1}$.  For points on different faces, the Euclidean distance is a lower bound for that length.

@@

To reduce the computations needed to calculate the energy score we define a kernel as {\bf Please write a clearer description here.} our kernel an upper bound on distance between points on separate faces as the length of the path from the source point, to an optimal point on the intersection between faces, to the destination point.
\begin{prop}\label{prop:g}
For points $a,b \in {\mathbb S}_{\infty}^{d-1}$ the kernel defined as  
  \begin{equation}
    g(\bm{a},\bm{b}) = \begin{cases}
        \pnorm{\bm{b}-\bm{a}}{2} &\text{ if }\argmax_{\ell}\bm{a} = \argmax_{\ell}\bm{b}\\
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} &\text{ otherwise}
    \end{cases} ,
  \end{equation}
  where $\bm{c}$ is in the intersection of the faces corresponding to $\bm{a}$ and $\bm{b}$, and minimizes $g(\bm{a},\bm{b})$, is negative definite.
\end{prop}

{\em Proof:}
When $\bm{a}$ and $\bm{b}$ are on the same face, then $g$ is the Euclidean distance, which is a 
  negative definite kernel.  When $\bm{a}$ and $\bm{b}$ are on separate
  faces, we have that $g(\bm{a}, \bm{b}) = g(\bm{b}, \bm{a})$ due to the uniqueness of $\bm{c}$ and the symmetry of the Euclidean distance. Additionally, given $n$ and  collection $\alpha_1,\ldots,\alpha_n\in{\mathbb R}$ such that  $\sum_{i = 1}^n\alpha_i = 0$, then
  \begin{equation*}
    \begin{aligned}
      \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j g(\bm{x}_1,\bm{x}_2) &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j \bigg[\pnorm{\bm{c} - \bm{x}_1}{2} + \pnorm{\bm{x}_2 - \bm{c}}{2}\bigg]\\
      &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{c} - \bm{x}_1}{2} + \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{x}_2 - \bm{c}}{2} \leq0,
    \end{aligned}
  \end{equation*}
  as the Euclidean norm is a negative definite kernel.$\hfill\Box$

Evaluating $g$ as defined in Proposition \ref{prop:g} requires the solution of a $d-2$-dimensional 
  optimization problem in order to obtain $\bm{c}$.  However, a computationally easier \makenote{rewrite sentence} method exists.
\begin{prop}
    Let $\bm{a} \in {\mathbb C}_{\ell}^{d-1}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-1}$.  The 
  rotation $P_{\jmath\ell}(\cdot)$ required to rotate the $\jmath$th face along the $\ell$th axis 
  produces the vector $\bm{b}^\prime$, where
  \begin{equation}
    \label{eqn:rotation}
     \bm{b}^{\prime}_i = P_{\jmath\ell}(\bm{b}) = 
    \begin{cases}
        b_{i} &\text{for }i\neq \jmath,\ell\\
        1 &\text{for }i = \ell\\
        2 - b_{\ell} &\text{for }i = \jmath
    \end{cases}.
  \end{equation}
  Then $g(\bm{a},\bm{b}) = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}$.
\end{prop}
{\em Proof:}
Note that for 
  $\bm{c} \in {\mathbb C}_{\jmath\ell}^{d-2} = {\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}$,
  $P_{\jmath\ell}(\bm{c}) = \bm{c}$.  Then
  \begin{equation*}
     \pnorm{\bm{b} - \bm{c}}{2} = \pnorm{P_{\jmath\ell}(\bm{b} - \bm{c})}{2} = \pnorm{\bm{b}^{\prime} - \bm{c}}{2},
  \end{equation*}
  as rotations preserve Euclidean norms.  Then,
  \begin{equation*}
    \begin{aligned}
    g(\bm{a}, \bm{b}) &= \pnorm{\bm{a} - \bm{c}}{2} + \pnorm{\bm{b} - \bm{c}}{2}\\
    &= \pnorm{\bm{a} - \bm{c}}{2} + \pnorm{\bm{b}^{\prime} - \bm{c}}{2} = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}
    \end{aligned}
  \end{equation*}
  when $\bm{c} \in \overrightarrow{\bm{a}\bm{b}^{\prime}}$, as the lower bound of the triangle inequality.  We have already declared $\bm{c}\in{\mathbb C}_{\jmath\ell}^{d-2}$, so it must reside at $\overrightarrow{\bm{a}\bm{b}^{\prime}}\cap{\mathbb C}_{\jmath\ell}^{d-2}$.  Because the straight line between $\bm{a}$ and $\bm{b}^{\prime}$ passes through this point without further analysis, the optimization of $\bm{c}$ is accomplished \emph{en~passant}.$\hfill\Box$

We report the average Interpretation of the energy score as we have formulated it here is the same as that of posterior predictive loss: smaller is better.

% EOF

>>>>

\subsection{Posterior predictive loss}
The posterior predictive loss criterion is introduced in \cite{gelfand1998}.  When we
  assume a squared error loss function, then for the $i$th observation, the posterior predictive
  loss criterion is computed as
  \begin{equation}
    \label{eq:ppl}
      S_{\text{PPL}}^k\left(P,x_i\right) = \text{Var}_P\left(X_i\right) + \frac{k}{k+1}\left(\text{E}_P\left[X_{i}\right] - x_i\right)^2
  \end{equation}
  where $X_i$ is a random variable from the posterior predictive distribution for $x_i$.  The
  scalar $k$ is a weighting factor that scales the importance of goodness of fit
  relative to precision.  In fact a small $\text{Var}(X_{i})$ indicates high
  precision, and a small $(\text{E}[X_{i}] - x_{i})^2$  indicates a good fit.  In our analysis,
  we take the limit as $k\to\infty$, weighting both
  equally. Overall, the smaller the score, the better.  Note that the PPLC score is defined 
  for a univariate distribution.  As we are dealing
  with a multivariate distribution, we will be taking the posterior predictive loss
  summed over all dimensions.  That is,
  \begin{equation}
    \label{eq:ppl2}
      S_{\text{PPL}}\left(P,\bm{x}_i\right) = \frac{1}{d}\sum_{\ell = 1}^d\left[\text{Var}_P\left(X_{i\ell}\right) + \left(\text{E}_P\left[X_{i\ell}\right] - x_{i\ell}\right)^2\right] .
  \end{equation}
  We report the average $S_{\text{PPL}}$ taken from all observed data.

\subsection{Energy score}
We focus on \emph{energy scores} defined for a general probability distribution $P$, with finite expectation, as 
    \begin{equation}
    \label{eq:es}
    \text{S}_{\text{ES}}\left(P, \bm{x}_i\right) =  \text{E}_p g\left(\bm{X}_i, \bm{x}_i\right) -
                                    \frac{1}{2}\text{E}_p g\left(\bm{X}_i,\bm{X}_i^{\prime}\right),
    \end{equation}
where $g$ is a kernel function. The score defined in Equation \eqref{eq:es} can be evaluated
  using samples from $P$, with the help of the law of large numbers.
  Moreover, Theorem~4 in \cite{gneiting2007}, states that if $g(\cdot,\cdot)$ is a negative 
  definite kernel, then $\text{S}(P,\bm{x})$ is a \emph{proper} scoring rule.  Recall that a 
  real valued function $g$ is a negative definite kernel if it is symmetric in its arguments, 
  and $\sum_{i=1}^n\sum_{j=1}^na_ia_jg(x_i,x_j)\leq 0$ for all positive integers $n$, and any 
  collection $a_1,\ldots,a_n\in{\mathbb R}$ such that  $\sum_{i = 1}^na_i = 0$.

In a Euclidean space, these conditions are satisfied by the Euclidean distance. However, in
  ${\mathbb S}_{\infty}^{d-1}$, Euclidean distance will tend to under-estimate the actual 
  distance required to travel between two points.  Let
\begin{equation*}
    {\mathbb C}_{\ell}^{d-1} = \lbrace \bm{x} : \bm{x} \in {\mathbb S}_{\infty}^{d-1}, x_{\ell} = 1\rbrace
\end{equation*}
comprise the $\ell$th \emph{face}.  For points on the same face, Euclidean distance 
  corresponds to the length of the shortest possible path in ${\mathbb S}_{\infty}^{d-1}$.  
  For points on different faces, the Euclidean distance is a lower bound for that length.

@@

To reduce the computations needed to calculate the energy score we define a kernel as {\bf Please write a clearer description here.} our kernel an upper bound on distance between points on separate faces as the length of the path from the source point, to an optimal point on the intersection between faces, to the destination point.
\begin{prop}\label{prop:g}
Let $\bm{a} \in {\mathbb C}_{\ell}^{d-1}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-1}$, for $\ell, \jmath \in \{1, \ldots , d\}$. Define $g$ as
\[  
    g(\bm{a},\bm{b}) = \min_{\bm{c} \in{\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}}\left\{ 
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} \right\}.
\]
Then $g$  is negative definite kernel.
\end{prop}

{\em Proof:}
When $\bm{a}$ and $\bm{b}$ are on the same face, then $\ell = \jmath$, and $g$ is the Euclidean distance between $\bm{a}$ and $\bm{b}$, which is a  negative definite kernel.  When $\bm{a}$ and $\bm{b}$ are on separate
  faces, let 
  \[\bm{d} = 
    \argmin_{\bm{c} \in{\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}}\left\{ 
        \pnorm{\bm{c}-\bm{a}}{2} + \pnorm{\bm{b}-\bm{c}}{2} \right\}.
  \]
  We have that $g(\bm{a}, \bm{b}) = g(\bm{b}, \bm{a})$ due to the uniqueness of $\bm{d}$ and the symmetry of the Euclidean distance. Additionally, given $n$ and  a collection $\alpha_1,\ldots,\alpha_n\in{\mathbb R}$ such that  $\sum_{i = 1}^n\alpha_i = 0$, then, 
  \begin{equation*}
    \begin{aligned}
      \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j g(\bm{x}_1,\bm{x}_2) &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j \bigg[\pnorm{\bm{d} - \bm{x}_1}{2} + \pnorm{\bm{x}_2 - \bm{d}}{2}\bigg]\\
      &= \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{d} - \bm{x}_1}{2} + \sum_{i = 1}^n\sum_{j = 1}^n \alpha_i\alpha_j\pnorm{\bm{x}_2 - \bm{d}}{2} \leq0,
    \end{aligned}
  \end{equation*}
  as the Euclidean norm is a negative definite kernel.$\hfill\Box$

Evaluating $g$ as defined in Proposition \ref{prop:g} requires the solution of a $(d-2)$-dimensional 
  optimization problem.  The following proposition provides a straightforward solution.
\begin{prop}
    Let $\bm{a} \in {\mathbb C}_{\ell}^{d-1}$, and $\bm{b} \in {\mathbb C}_{\jmath}^{d-1}$, for $\ell, \jmath \in \{1, \ldots , d\}$.  For $\ell\neq \jmath$, the 
  transformation $P_{\jmath\ell}(\cdot)$ required to rotate the $\jmath$th face along the $\ell$th axis 
  produces the vector $\bm{b}^\prime$, where
  \begin{equation}
    \label{eqn:rotation}
     \bm{b}^{\prime}_i = P_{\jmath\ell}(\bm{b}) = 
    \begin{cases}
        b_{i} &\text{for }i\neq \jmath,\ell\\
        1 &\text{for }i = \ell\\
        2 - b_{\ell} &\text{for }i = \jmath
    \end{cases}.
  \end{equation}
  Then $g(\bm{a},\bm{b}) = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}$.
\end{prop}
{\em Proof:}
Notice that for 
  $\bm{c} \in {\mathbb C}_{\jmath}^{d-1}\cap{\mathbb C}_{\ell}^{d-1}$,
     $\pnorm{\bm{b} - \bm{c}}{2} =  \pnorm{\bm{b}^{\prime} - \bm{c}}{2}$. {\bf This is true just because you do the math, the invariance of Euclidean norms is not needed.} Then, recalling the definition of $\bm{d}$ in the proof of Proposition \ref{prop:g}, we have that
  \begin{equation*}
    \begin{aligned}
    g(\bm{a}, \bm{b}) &= \pnorm{\bm{a} - \bm{d}}{2} + \pnorm{\bm{b} - \bm{d}}{2}\\
    &= \pnorm{\bm{a} - \bm{d}}{2} + \pnorm{\bm{b}^{\prime} - \bm{d}}{2} = \pnorm{\bm{a} - \bm{b}^{\prime}}{2}.
    \end{aligned}
  \end{equation*}
  The last equality is due to the fact that $\bm{a}$ and $\bm{b}'$ belong to the same hyperplane.
  $\hfill\Box$

We report the average $S_{\text{ES}}$ taken across all observed data.  In this form, the interpretation of $S_{\text{ES}}$ is similar to that of $S_{\text{PPL}}$: the smaller the score, the better.
% EOF

### 2021-11-11 methodology_projgamma.tex
updates

A key feature of the the projected Gamma distribution is its computational properties. We augment  $\mathcal{PG}(\bm{y}_i\mid\bm{\alpha}_i,\bm{\beta}_i) $ by introducing a latent radial component $r_i$, for each observation. Using Equation \eqref{joint} we observe that the 
full conditional of $r_i$ is easy to sample from, as it is given as
\begin{equation}
    r_i\mid\bm{\alpha}_i,\bm{\beta}_i \sim \mathcal{G}\left(\sum_{\ell = 1}^d\alpha_{i\ell}, \sum_{\ell = 1}^d\beta_{\ell} y_{i\ell}\right).
\end{equation}
Moreover,  the full conditional for $\bm{\alpha}_j,\bm{\beta}_j$ is then proportional to

\begin{equation}
    f(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{Y},\bm{r},\bm{\delta},\ldots) \propto \prod_{i:\gamma_i = j}\prod_{\ell = 1}^d\mathcal{G}\left(r_iy_{i\ell}\mid\alpha_{j\ell},\beta_{j\ell}\right) \times dG_0(\bm{\alpha}_j,\bm{\beta}_j)
\end{equation}

% \begin{equation}
%     f(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{Y},\bm{r},\bm{\delta},\ldots) \propto \prod_{i:\gamma_i = j}\prod_{\ell = 1}^d\mathcal{G}\left(r_iy_{i\ell}\mid\alpha_{j\ell},\beta_{j\ell}\right) \times dG_0(\alpha_{\ell},\beta_{\ell})
% \end{equation}
% {\bf I am confused here. Shouldn't it be: }

% \begin{equation}
%     f(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{Y},\bm{r},\bm{\delta},\ldots) \propto \prod_{i:\gamma_i = j}\left(\prod_{\ell = 1}^d\mathcal{G}\left(r_iy_{i\ell}\mid\alpha_{j\ell},\beta_{j\ell}\right)\right) \times dG_0(\bm{\alpha}_{i},\bm{\beta}_{i})
% \end{equation}
We first consider a centering distribution given by a product of independent Gammas:
\begin{equation*}
    G_0(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{\xi},\bm{\tau},\bm{\zeta},\bm{\sigma}) = \prod_{\ell = 1}^d\mathcal{G}(\alpha_{j\ell}\mid \xi_{\ell},\tau_{\ell})\times\prod_{\ell = 2}^d\mathcal{G}(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell}).
\end{equation*}
This model is completed with independent Gamma priors on $\xi_{\ell}$, $\tau_{\ell}$, $\zeta_{\ell}$, $\sigma_{\ell}$.  We also assume a Gamma prior on $\eta$, that is updated via the procedure outlined in \cite{escobar1995}.  We refer to this model as the \emph{projected Gamma--Gamma} (PG--G) model.  An advantage of the PG--G model is that, thanks to conjugacy, the rate parameters can easily be integrated out for inference on $\bm{\alpha}_j$.  Then, the full conditional for $\alpha_{j\ell}$ takes the form
\begin{equation*}
    \pi(\alpha_{j\ell}\mid \bm{r},\bm{Y},\bm{\gamma},\xi_\ell,\tau_\ell,\zeta_\ell,\sigma_\ell) \propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i\ell}\right)^{\alpha_{j\ell} - 1}\alpha_{j\ell}^{\xi_\ell - 1}e^{-\tau_\ell \alpha_{j\ell}}\times\Gamma\left(n_j\alpha_{j\ell} + \zeta_{\ell}\right)}{\Gamma^{n_j}(\alpha_{j\ell})\left(\sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right)^{n_j\alpha_{j\ell} + \zeta_{\ell}}}
\end{equation*}
for $\ell = 2,\ldots,d$.  For $\ell = 1$, as $\beta_{1} := 1$, the full conditional takes the form
\begin{equation}
    \label{eqn:alpha1update}
    \pi(\alpha_{j1}\mid\bm{r},\bm{Y},\bm{\gamma},\xi_1,\tau_1) \propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i1}\right)^{\alpha_{j1} - 1}\alpha_{j1}^{\zeta_1 - 1}e^{-\tau_1\alpha_{j1}}}{\Gamma^{n_j}(\alpha_{j1})}.
\end{equation}
Sample of $\alpha_{j\ell}$ can thus be obtained using a Metropolis step. In our analysis, we first transform $\alpha_{j\ell}$ to the log scale, and use a normal proposal density.
The full conditional for $\beta$ is 
\begin{equation}
    \beta_{j\ell}\mid\bm{r},\bm{Y},\alpha,\zeta_{\ell},\sigma_{\ell} \sim \mathcal{G}\left(\beta_{j\ell}\mid n_j\alpha_{j\ell} + \zeta_\ell, \sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right),
\end{equation}
for $\ell = 2,\ldots, d$.  Updating $\beta_{j\ell}$ is done via a Gibbs step.  The hyper-parameters $\xi_{\ell},\tau_{\ell},\zeta_{\ell},\sigma_{\ell}$ follow similar Gamma-Gamma update relationships.  We also explore a restricted form of this model, where $\beta_{\ell} = 1$ for all $\ell$.  Under this model, we use the full conditional in Equation~\ref{eqn:alpha1update} for all $\ell$, and omit inference on $\bm{\zeta},\bm{\sigma}$.  We refer to this model as the \emph{projected restricted Gamma--Gamma} (PRG--G) model.

The second form of centering distribution we explore is a multivariate log-normal distribution on the shape parameters $\bm{\alpha}_j$, with independent Gamma $\beta_{j\ell}$ rate parameters.  
\begin{equation}
    G_0\left(\bm{\alpha}_j,\bm{\beta}_j\mid\bm{\mu},\bm{\Sigma},\zeta,\sigma\right) = \mathcal{LN}\left(\bm{\alpha}_j\mid\bm{\mu},\bm{\Sigma}\right)\times\prod_{\ell = 2}^d\mathcal{G}\left(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell}\right).
\end{equation}
This model is completed with a normal prior on $\bm{\mu}$, an inverse Wishart prior on $\bm{\Sigma}$, and Gamma priors on $\zeta_{\ell}$ and $\sigma_{\ell}$, and $\eta$.  This model is denoted as the \emph{projected Gamma--log-normal} (PG--LN) model.  We also explore a restricted Gamma form of this model as above, where $\beta_{\ell} = 1$ for all $\ell$.  This is denoted as the \emph{projected restricted Gamma--log-normal} (PRG--LN) model.  In the interest of brevity, we omit the inferential forms for this model, sufficing to say that updating $\bm{\alpha}_j$ is accomplished with a joint Metropolis-within-Gibbs step; the normal prior for $\mu$ is conjugate for the log-normal $\log\bm{\alpha}_j$, and updated via a Gibbs step; and an inverse Wishart prior for $\Sigma$ is again conjugate to the log-normal $\bm{\alpha}_j$, implying that it can also be updated via a Gibbs step.

>>>>

A key feature of the the projected Gamma distribution is its computational properties. We augment  $\mathcal{PG}(\bm{y}_i\mid\bm{\alpha}_i,\bm{\beta}_i) $ by introducing a latent radial component $r_i$, for each observation. Using Equation \eqref{joint} we observe that the 
full conditional of $r_i$ is easy to sample from, as it is given as
\begin{equation}
    r_i\mid\bm{\alpha}_i,\bm{\beta}_i \sim \mathcal{G}\left(\sum_{\ell = 1}^d\alpha_{i\ell}, \sum_{\ell = 1}^d\beta_{\ell} y_{i\ell}\right).
\end{equation}
Moreover,  the full conditional for $\bm{\alpha}_j,\bm{\beta}_j$ is then proportional to
\begin{equation}
    f(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{Y},\bm{r},\bm{\delta},\ldots) \propto \prod_{i:\gamma_i = j}\prod_{\ell = 1}^d\mathcal{G}\left(r_iy_{i\ell}\mid\alpha_{j\ell},\beta_{j\ell}\right) \times dG_0(\bm{\alpha}_j,\bm{\beta}_j).
\end{equation}
We first consider a centering distribution given by a product of independent Gammas:
\begin{equation*}
    G_0(\bm{\alpha}_j,\bm{\beta}_j\mid \bm{\xi},\bm{\tau},\bm{\zeta},\bm{\sigma}) = \prod_{\ell = 1}^d\mathcal{G}(\alpha_{j\ell}\mid \xi_{\ell},\tau_{\ell})\times\prod_{\ell = 2}^d\mathcal{G}(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell}).
\end{equation*}
This model is completed with independent Gamma priors on $\xi_{\ell}$, $\tau_{\ell}$, $\zeta_{\ell}$, $\sigma_{\ell}$.  We also assume a Gamma prior on $\eta$, that is updated via the procedure outlined in \cite{escobar1995}.  We refer to this model as the \emph{projected Gamma--Gamma} (PG--G) model.  An advantage of the PG--G model is that, thanks to conjugacy, the rate parameters can easily be integrated out for inference on $\bm{\alpha}_j$.  Then, the full conditional for $\alpha_{j\ell}$ takes the form
\begin{equation*}
    \pi(\alpha_{j\ell}\mid \bm{r},\bm{Y},\bm{\gamma},\xi_\ell,\tau_\ell,\zeta_\ell,\sigma_\ell) \propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i\ell}\right)^{\alpha_{j\ell} - 1}\alpha_{j\ell}^{\xi_\ell - 1}e^{-\tau_\ell \alpha_{j\ell}}\times\Gamma\left(n_j\alpha_{j\ell} + \zeta_{\ell}\right)}{\Gamma^{n_j}(\alpha_{j\ell})\left(\sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right)^{n_j\alpha_{j\ell} + \zeta_{\ell}}}
\end{equation*}
for $\ell = 2,\ldots,d$.  For $\ell = 1$, as $\beta_{1} := 1$, the full conditional takes the form
\begin{equation}
    \label{eqn:alpha1update}
    \pi(\alpha_{j1}\mid\bm{r},\bm{Y},\bm{\gamma},\xi_1,\tau_1) \propto \frac{\left(\prod_{i:\gamma_i = j}r_iy_{i1}\right)^{\alpha_{j1} - 1}\alpha_{j1}^{\zeta_1 - 1}e^{-\tau_1\alpha_{j1}}}{\Gamma^{n_j}(\alpha_{j1})}.
\end{equation}
Sample of $\alpha_{j\ell}$ can thus be obtained using a Metropolis step. In our analysis, we first transform $\alpha_{j\ell}$ to the log scale, and use a normal proposal density.
The full conditional for $\beta$ is 
\begin{equation}
    \beta_{j\ell}\mid\bm{r},\bm{Y},\alpha,\zeta_{\ell},\sigma_{\ell} \sim \mathcal{G}\left(\beta_{j\ell}\mid n_j\alpha_{j\ell} + \zeta_\ell, \sum_{i:\gamma_i = j}r_iy_{i\ell} + \sigma_{\ell}\right),
\end{equation}
for $\ell = 2,\ldots, d$.  Updating $\beta_{j\ell}$ is done via a Gibbs step.  The hyper-parameters $\xi_{\ell},\tau_{\ell},\zeta_{\ell},\sigma_{\ell}$ follow similar Gamma-Gamma update relationships.  We also explore a restricted form of this model, where $\beta_{\ell} = 1$ for all $\ell$.  Under this model, we use the full conditional in Equation~\ref{eqn:alpha1update} for all $\ell$, and omit inference on $\bm{\zeta},\bm{\sigma}$.  We refer to this model as the \emph{projected restricted Gamma--Gamma} (PRG--G) model.

The second form of centering distribution we explore is a multivariate log-normal distribution on the shape parameters $\bm{\alpha}_j$, with independent Gamma $\beta_{j\ell}$ rate parameters.  
\begin{equation}
    G_0\left(\bm{\alpha}_j,\bm{\beta}_j\mid\bm{\mu},\bm{\Sigma},\zeta,\sigma\right) = \mathcal{LN}\left(\bm{\alpha}_j\mid\bm{\mu},\bm{\Sigma}\right)\times\prod_{\ell = 2}^d\mathcal{G}\left(\beta_{j\ell}\mid\zeta_{\ell},\sigma_{\ell}\right).
\end{equation}
This model is completed with a normal prior on $\bm{\mu}$, an inverse Wishart prior on $\bm{\Sigma}$, and Gamma priors on $\zeta_{\ell}$ and $\sigma_{\ell}$, and $\eta$.  This model is denoted as the \emph{projected Gamma--log-normal} (PG--LN) model.  We also explore a restricted Gamma form of this model as above, where $\beta_{\ell} = 1$ for all $\ell$.  This is denoted as the \emph{projected restricted Gamma--log-normal} (PRG--LN) model.  In the interest of brevity, we omit the inferential forms for this model, sufficing to say that the Gamma prior for $\beta_{j\ell}$ means it can again be integrated out, improving mixing on sampling $\bm{\alpha}_j$ with a joint Metropolis step; the normal prior for $\mu$ is conjugate for the log-normal $\log\bm{\alpha}_j$, and updated via a Gibbs step; and an inverse Wishart prior for $\Sigma$ is again conjugate to the log-normal $\bm{\alpha}_j$, implying that it can also be updated via a Gibbs step.

### 2021-11-11 results.tex
updates

\begin{figure}[ht!]
    \centering
    \caption{Pairwise extremal dependence coefficients for IVT data, from fitted PRG--G.  
    The left corresponds to the 8-cell data, the right the 47-cell data.\label{fig:chi_ij}}
    \begin{minipage}{.49\textwidth}
      \centering
      \includegraphics[width=0.99\linewidth]{./images/chi_ij_8}
    \end{minipage}
    \begin{minipage}{.49\textwidth}
      \centering
      \includegraphics[width=0.99\linewidth]{./images/chi_ij_46}
    \end{minipage}
\end{figure}

@@

% EOF

>>>>

\begin{figure}[ht]
    \centering
    \caption{Pairwise extremal dependence coefficients for IVT data, from fitted PRG--G.  
    The left corresponds to the 8-cell data, the right the 47-cell data.\label{fig:chi_ij}}
    \begin{minipage}{.49\textwidth}
      \centering
      \includegraphics[width=0.99\linewidth]{./images/chi_ij_8}
    \end{minipage}
    \begin{minipage}{.49\textwidth}
      \centering
      \includegraphics[width=0.99\linewidth]{./images/chi_ij_46}
    \end{minipage}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \caption{Conditional survival curves for selected dimensions, conditioning on all other dimensions at greater than 90th percentile (fitted)\label{fig:condsurv1d}}
    \includegraphics{./images/condsurv_1d}
\end{figure}

\begin{figure}[htb]
    \centering
    \caption{Pairwise conditional survival curves for selected dimensions, conditioning on all other dimensions at greater than 90th percentile (fitted)\label{fig:condsurv2d}}
    \includegraphics[height=5in, width=5in]{./images/condsurv_2d}
\end{figure}

In Figure~\ref{fig:condsurv1d}, using the PRG--G model on the 8-dimensional data, we implement 
    the conditional survival curve defined in Equation~\ref{eqn:condsurv}, for one dimension, 
    conditioned on all other dimensions being greater than their (fitted) $90$th percentile.  
    On the left, the survival curve is in original units, whereas the right is the same plot 
    given in standardized units.  
    Figure~\ref{fig:condsurv2d} implements the same conditional survival equation in pairwise form,
    conditioning on all other dimensions.  Here we see another representation of extremal dependence,
    affecting the shape of the conditional survival curves.  The top pair represents the joint survival function between grid zones 1 and 2, which we saw from Figure~\ref{fig:chi_ij} to exhibit strong
    extremal dependence.  We see that again here, where the joint survival surface in standardized form
    is strongly convex.  The bottom pair represents the joint survival surface between grid zones 2 and 8, which exhibited low extremal dependence.  We see the effects of that, in the concave pairwise survival surface.
% EOF
